\documentclass[../MA_Thesis.tex]{subfiles}
\renewcommand{\baselinestretch}{1.5} 
\usepackage{hyperref}
\usepackage{csquotes}

\begin{document}
Building on significant advancements in affective psychology and creativity research, this study acknowledges the well-documented synergy between mood states and creative cognition. As described in the literature, frameworks categorizing mood along pleasant-unpleasant and activated-deactivated dimensions have elucidated the intricate relationships between mood induction methods and their cognitive consequences (\cite{siedlecka_experimental_2019}). Coupled with the burgeoning exploration of creativity's multifaceted nature---encompassing definitions, underlying cognitive processes, and diverse assessment tasks (\cite{kaufman_cambridge_2010})---this mutual enrichment has paved the way for both theoretical propositions and empirical validations of the mood-creativity linkage.

Echoing \textcite{kaufman_cambridge_2010}'s call for innovative assessment techniques that more accurately reflect the dynamics of creative thinking, this study aims to advance the debate on the mood-creativity connection by introducing a novel \textit{task-measurement} combination, which leverages both the versatility of drawing tasks to uncover the cognitive underpinnings of creativity and employing state-of-the-art artificial intelligence techniques. Focusing on \textit{domain-general}, \textit{little-c} creativity, our study aims to capture the nuanced effects of mood on creative processes and, more specifically, the (potential) building effects of positive mood states on thought-action repertoires (\cite{fredrickson_role_2001}). By distinguishing positive mood states varying on the activation dimension---happiness (high activation) and calmness (low activation), this study seeks to scrutinize the hypothesized flexibility pathway (as suggested by \textcite{de_dreu_hedonic_2008}'s dual pathway to creativity model) in which positive activating mood states, rather than positive deactivating ones, predict cognitive flexibility, characterized by employing wide-ranging and comprehensive cognitive categories to form associations. The conducive influence of positive (activating) mood is further hypothesized to enhance the originality aspect of creativity (i.e., the uncommonness of ideas, solutions, or products). Specifically, this study is guided by the following two research questions:  
\begin{enumerate}
    \item \textbf{How do positive moods across the spectrum of activation level, including positive moods with high level of activation (e.g., happiness) and positive moods with low level of activation (e.g., calmness), affect cognitive flexibility during the creative ideation process, respectively?}
    \item \textbf{How does cognitive flexibility during the creative ideation process further influence the originality aspect of creativity in the final product (i.e., whether flexibility mediates the relationship between positive mood and the originality aspect of creativity)?}
\end{enumerate}

To answer these research questions, this study adopts a validated film-based mood induction protocol (\cite{siedlecka_experimental_2019}) for mood induction. We randomly assign participants to one of three mood conditions (high-arousal positive, low-arousal positive, or neutral) to explore the effects of positive mood states, ranging from activating to deactivating, on the flexibility pathway of creativity. To capture the flexibility and originality aspects of creativity, we utilize tasks that not only track the dynamics of creative processes, but also incorporate novel methodologies from generative sketch modeling and NLP to examine the proposed building effects of positive mood states on thought-action repertoires (\cite{fredrickson_role_2001}). 

Following \textcite{barbot_dynamics_2018}'s Multi-Trial Creative Ideation (MTCI) framework (featuring multi-stimuli approach \& dynamics of ideation process), we adopt the incomplete shape drawing task and narrative on creative ideation processes to record the final creative output, but also track the dynamics of creative thinking that leads to the final completed drawings. The collection of both behavioral (stroke-level drawing data) and semantic (narrative-based) data therefore allows us to capture the flexibility aspect of creativity through two complementart perspectives. First, the Compositional Stroke Embedding (CoSE) model, a generative model trained on human drawing data, uses a Gaussian Mixture Model (GMM) to predict stroke trajectories. Flexibility is quantified via average entropy and Bhattacharyya distance (capturing sustained exploratory breadth), as well as the inflection proportions (capturing adaptive strategy switching) throughout the drawing process. Second, Divergent Semantic Integration (DSI), a measure based on BERT embeddings, captures the conceptual distance between narrative elements to reflect semantic flexibility. Meanwhile, the originality aspect of creativity (i.e., the originality of the final completed drawings) is evaluated using AuDrA, a deep learning model trained on human-rated drawings from the same task. Mediation analyses are conducted to test whether flexibility mediates the effect of mood condition on originality. Mood condition is treated as a multicategorical independent variable, with the neutral group serving as the reference category. Dummy coding is used to create two binary contrasts: one comparing the high-arousal positive mood group to the neutral control (D1), and another comparing the low-arousal positive mood group to the neutral control (D2). Each flexibility metric is tested as a separate mediator in a parallel mediation framework.

The primary contribution of this study is its integration of drawing tasks with deep learning and natural language processing techniques to quantitatively assess both the flexibility and originality aspects of creativity. This method offers a more comprehensive analysis than previous studies that relied solely on verbal tasks such as the alternative use task and the remote association task (e.g., \cite{kenett_investigating_2014}; \cite{kenett_flexibility_2018}) to quantify the flexibility aspect of creativity. Complementing the predominance of verbal creativity assessments with visual creativity (a canonical form of creative expression; \cite{morrisskay_evolution_2010}) allows for a holistic understanding of creative processes. Verbal and visual creativity engage different cognitive and neural pathways; verbal creativity often relies on language-based processes and abstract thinking (\cite{benedek_create_2014}), while visual creativity engages spatial reasoning and visual-motor coordination (\cite{schlegel_artist_2015}). By studying both aspects, this proposed study could uncover complementary insights into the cognitive mechanisms underlying the flexibility aspect of creativity. This integrative approach enhances the validity of the measurements by capturing a broader range of creative expression and providing a more nuanced understanding of the dynamics of creative thinking. Meanwhile, when it comes to assessing the originality aspect of creativity (especially visual creativity), using automated scoring methods effectively addresses several practical limitations in creativity research. These include the high labor costs associated with manual evaluations and the inherent subjectivity that can bias expert ratings (\cite{patterson_audra_2023}). Automated methods provide a scalable and consistent way to evaluate originality, ensuring that the assessments are objective and replicable across different studies.

\end{document}